Reasoning about a Capability Machine with Local Capabilities\\ Provably Safe Stack and Return Pointer Management (without OS Support)

### 2017-03-15 csf-2017/paper.tex
inserted authors and ack., so it is ready for distribution
typos

% \author{%
% \IEEEauthorblockN{Lau~Skorstengaard}
% \IEEEauthorblockA{Aarhus University, Denmark\\
% Email: lau@cs.au.dk} \and
% \IEEEauthorblockN{Dominique~Devriese}
%   \IEEEauthorblockA{iMinds-DistriNet, KU Leuven, Belgium\\
% Email: dominique.devriese@cs.kuleuven.be} \and
% \IEEEauthorblockN{Lars~Birkedal}\IEEEauthorblockA{Aarhus University, Denmark\\
% Email: birkedal@cs.au.dk}}
\author{}
\maketitle

@@

Note that storing stack pointers for use during future invocations would also be
dangerous in itself, i.e. not just because it can be used to store return
pointers. Imagine the adversary stores his stack pointer, invokes trusted
code that uses part of the stack to store private data and then invokes the
adversary again with a stack pointer restricted to exclude the part containing
the private data. If the adversary had a way of keeping hold of their old stack
pointer, it could access the private data stored there by the trusted code, and
break local-state encapsulation.

@@

Our answer is that the adversary should provide it to us, much like we provide
them with a stack pointer when we invoke their code. However, it is important
that we do not just accept any capability as a stack pointer but check that it
is safe to use. Specifically, we check that it is indeed a $\rwlx$ capability. Without
this check, an adversary could potentially get control over our local stack
frame during a callback by passing us a local $\rwx$ capability to a global data
structure instead of a proper stack pointer
and a global callback for our callback to invoke. If our local state contains no
local capabilities, then, following our calling convention, the callback would
not fail and the adversary could use a stored capability for the global data
structure to access our local state. To prevent this from happening, we need to
make sure the stack capability gives $\rwlx$ authority, since the system wide assumption then
tells us that the adversary cannot have a global capabilities to our local stack.

@@

\begin{comment}
\section*{Acknowledgements}
\label{sec:acknowledgements}

This research was supported in part by the ModuRes Sapere Aude Advanced Grant from The Danish Council for Independent Research for the Natural Sciences (FNU).
Dominique Devriese holds a postdoctoral fellowship from the Research Foundation - Flanders (FWO).
\end{comment}

>>>>

\author{%
  % \IEEEauthorblockN{Lau~Skorstengaard}
  % \IEEEauthorblockA{Aarhus University, Denmark\\
  %   Email: lau@cs.au.dk} \and
  % \IEEEauthorblockN{Dominique~Devriese}
  % \IEEEauthorblockA{iMinds-DistriNet, KU Leuven, Belgium\\
  %   Email: dominique.devriese@cs.kuleuven.be} \and
  % \IEEEauthorblockN{Lars~Birkedal}\IEEEauthorblockA{Aarhus University, Denmark\\
  %   Email: birkedal@cs.au.dk}
}
\maketitle

@@

Note that storing stack pointers for use during future invocations would also be
dangerous in itself, i.e. not just because it can be used to store return
pointers. Imagine the adversary stores their stack pointer, invokes trusted
code that uses part of the stack to store private data and then invokes the
adversary again with a stack pointer restricted to exclude the part containing
the private data. If the adversary had a way of keeping hold of their old stack
pointer, it could access the private data stored there by the trusted code and
break local-state encapsulation.

@@

Our answer is that the adversary should provide it to us, much like we provide
them with a stack pointer when we invoke their code. However, it is important
that we do not just accept any capability as a stack pointer but check that it
is safe to use. Specifically, we check that it is indeed a $\rwlx$ capability. Without
this check, an adversary could potentially get control over our local stack
frame during a callback by passing us a local $\rwx$ capability to a global data
structure instead of a proper stack pointer
and a global callback for our callback to invoke. If our local state contains no
local capabilities, then, following our calling convention, the callback would
not fail and the adversary could use a stored capability for the global data
structure to access our local state. To prevent this from happening, we need to
make sure the stack capability gives $\rwlx$ authority, since the system wide assumption then
tells us that the adversary cannot have global capabilities to our local stack.

@@

\section*{Acknowledgements}
\label{sec:acknowledgements}

This research was supported in part by the ModuRes Sapere Aude Advanced Grant from The Danish Council for Independent Research for the Natural Sciences (FNU).
Dominique Devriese holds a postdoctoral fellowship from the Research Foundation - Flanders (FWO).
